\begin{acknowledgements}

In the first place, I would like to dedicate my first words of gratitude to my supervisors, Andrea and Stefano.
I cannot thank you enough your guidance during these 4 years.
I will not be able to pay you back for the effort you have made to turn a stubborn person into someone flexible who can face the problems of the world of research.
The transformation that I have undergone during these four years is incredible.
Everything I have learned from you, both inside and outside the lab, is a gift.
Now this stage of my life is coming to an end, but I hope to continue working with you for many years because you have a lot of science inside you are wonderful people.
I couldn't have had better supervisors.

My gratitude also extends to Marco, because you were the person who taught me to work in a laboratory.
The endless hours in the lab turning the laser on and off must have had a serious effect on me.
I just hope that one day we can remove the panels article from our to-do list and we can move to new projects.
You should also know that the difficulty of my thesis would have been much greater if I had not had you there with your experience in the water tunnel and to deal with Stefano and, above all, with Andrea.
Nor can I forget the other members of the rest of the colleagues from the Experimental Aerodynamics and Propulsion Lab, especially Carlos and Rodrigo.
Without the beers when we left the laboratory, I would not have lasted until here.

I would like to thank the people from KTH.
Thanks to Ricardo, Hossein and Philipp ,for letting me participate in their projects and giving me the opportunity to use their resources.
Luca deserves a special mention, with whom I have been able to work side by side these last two years.
I wish you the best for the future.

I cannot forget the Bioengineering and Aerospace engineering department's professors.
I would like to mention especially Manolo, Oscar, Manuel Sanjurjo and Manuel Soler.
I have learnt a lot from you, thank you very much.

A mis amigos Andrea, Andrés, Dani, Gonzalo, Joaquín, Marta, Nere, Proven, Salme, Sergio, Sisa y Tamara, compañeros de viaje durante todos estos años de estudio, les reservo un lugar especial.
Las horas interminables en la biblioteca o en las salas de ordenadores del edificio siete al final tuvieron su recompensa.
A los compañeros de Boeing, con especial cariño para Javi, Pablo, Miguel y David, que me dieron la oportunidad de poder trabajar con ellos, y a Quique, que ha mantenido la confianza en mí.

A toda mi familia, de la que siempre he recibido un inmenso cariño.
Para mi, no hay mejores personas de las que estar rodeado, os quiero.
A Sandra, por ser mi apoyo durante estos últimos años.
Espero poder estar a tu lado cuando me necesites de la misma forma que has estado tu ahí para mi.
Te quiero.

Finalmente, a papá y mamá.
Los logros de esta tesis son más vuestros que míos.
No hay palabras que hagan justicia al gran esfuerzo que habéis hecho todos estos años para educarme, formarme y, sobre todo, quererme. Muchísimas gracias, os quiero.

\end{acknowledgements}
