%------------------------------------------------------------------------------
% Define title, author(s), affiliation and publishing status
%
\papertitle[Wall-bounded turbulence from wall quantities] % Short title used in healines (optional)
{%
Convolutional-network models to predict wall-bounded turbulence from wall quantities% THE COMMENT SYMBOL AT THE END OF THIS LINE IS NEEDED
}%
%
\papertoctitle{Wall-bounded turbulence from wall quantities} % Title for toc
%
\paperauthor[L. Guastoni \etal] % Short authors used in headlines and List Of Papers
{%
  Luca Guastoni$^{1,2}$, Alejandro Güemes$^3$, Andrea Ianiro$^3$, Stefano Discetti$^3$, Phillip Schlatter$^{1,2}$, Hossein Azizpour$^{4,2}$ and Ricardo Vinuesa$^{1,2}$%
}%
%
\listpaperauthor{Guastoni, L., Güemes, A., Ianiro, A., Discetti, S., Schlatter, P., Azizpour, H. \& Vinuesa, R.}% (optional) Short authors used in List Of Papers
%
\paperaffiliation
{%
  $^1$ Linn\'e FLOW Centre, KTH Mechanics, S-100 44 Stockholm, Sweden\\%
  $^2$ Swedish e-Science Research Centre (SeRC), Stockholm, Sweden\\%
  $^3$ Aerospace Research Group, Universidad Carlos III de Madrid, Leganés, Spain\\%
  $^4$ Division of Robotics, Perception and Learning, KTH Royal Institute of Technology, Stockholm, Sweeden\\
}%
%
\paperjournal[J. Fluid Mech., \textbf{Under review}] % Short publish info used in List Of Papers
{%
	Journal of Fluid Mechanics, \textbf{Under review}%
}%
%
\papervolume{}%
%
%\papernumber{1}
%
\paperpages{}%
%
\paperyear{2021}%
%
\papersummary%
{% Insert summary of the paper here (used in introduction)
Two models based on convolutional neural networks are trained to predict the two-dimensional instantaneous velocity-fluctuation fields at different wall-normal locations in a turbulent open-channel flow, using the wall-shear-stress components and the wall pressure as inputs.
The first model is a fully-convolutional neural network (FCN) which directly predicts the fluctuations, while the second one reconstructs the flow fields using a linear combination of orthonormal basis functions, obtained through proper orthogonal decomposition (POD), hence named FCN-POD.
Both models are trained using data from direct numerical simulations (DNS) at friction Reynolds numbers $Re_{\tau} = 180$ and 550.
Being able to predict the nonlinear interactions in the flow, both models show better predictions than the extended proper orthogonal decomposition (EPOD), which establishes a linear relation between input and output fields.
The performance of the models is compared based on predictions of the instantaneous fluctuation fields, turbulence statistics and power-spectral densities.
FCN exhibits the best predictions closer to the wall, whereas FCN-POD provides better predictions at larger wall-normal distances.
We also assessed the feasibility of transfer learning for the FCN model, using the model parameters learned from $Re_{\tau}=180$ dataset to initialize those of the model that is trained on the $Re_{\tau}=550$ dataset.
After training the initialized model at the new $Re_{\tau}$, our results indicate the possibility to match the reference-model performance up to $y^{+}=50$, with $50\%$ and $25\%$ of the original training data.
We expect that these non-intrusive sensing models will play an important role in applications related to closed-loop control of wall-bounded turbulence.
}
%
\graphicspath{{paper2/}}%
%
%
%===============================================================================
%                            BEGIN PAPER
%===============================================================================
%
\begin{paper}

\makepapertitle

%------------------------------------------------------------------------------
% Abstract
%------------------------------------------------------------------------------
%
\begin{paperabstract}
	Two models based on convolutional neural networks are trained to predict the two-dimensional velocity-fluctuation fields at different wall-normal locations in a turbulent-open channel flow, using the wall-shear-stress components and the wall pressure as inputs.
	The first model is a fully-convolutional neural network (FCN) which directly predicts the fluctuations, while the second one reconstructs the flow fields using a linear combination of orthonormal basis functions, obtained through proper orthogonal decomposition(POD), hence named FCN-POD.
	Both models are trained using data from two direct numerical simulations (DNS) at friction Reynolds numbers $Re_{\tau} = 180$ and 550.
	Thanks to their ability to predict the nonlinear interactions in the flow, both models show a better prediction  performance  than  the  extended  proper  orthogonal  decomposition  (EPOD), which establishes a linear relation between input and output fields.
	The performance of the various models is compared based on predictions of the instantaneous fluctuation fields, turbulence statistics and power-spectral densities.
	The FCN exhibits the best predictions closer to the wall, whereas the FCN-POD model provides better predictions at larger wall-normal distances.
	We also assessed the feasibility of performing transfer learning for the FCN model, using the weights from $Re_{\tau} = 180$ to initialize those of the $Re_{\tau} = 550$ case. Our results indicate that it is possible to obtain a performance similar to that of the reference model up to $y^+= 50$, with 50\% and 25\% of the original training data.
	These non-intrusive sensing models will play an important role in applications related to closed-loop control of wall-bounded turbulence
        \keywords{turbulent flows, turbulence simulation}
\end{paperabstract}


%------------------------------------------------------------------------------
% Article
%------------------------------------------------------------------------------
%
\input{paper2/article.tex}


%------------------------------------------------------------------------------
% Bibliography
%------------------------------------------------------------------------------
%
%\clearpage
\bibliographystyle{jfm}
\bibliography{thesis}
%
\IfFileExists{paper2/paper.bbl}{\input{paper2/paper.bbl}}{}


%===============================================================================
%                            END PAPER
%===============================================================================
\end{paper}
